# Samsung phone Data Analysis

## Preliminaries

Download the data and save as zip file
```{r}
download.file("http://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip","dataset.zip")

```
Unarchive the saved zip file
```{r}
unzip("dataset.zip")
```
## Cleaning data

Read variable names in feature.txt file
```{r}
features<-read.table("UCI HAR Dataset/features.txt")
```
Look at the data
```{r}
head(features)
```
Convert variable names to appropriate labels by using gsub function
```{r}
feature_names<-gsub("[\\(\\)\\,\\-]", "", features$V2)
```
Create index of columns with "mean" and "std" substrings
```{r}
mean.and.std<-grep("mean|std", feature_names, ignore.case=TRUE)
```
Read activity_labels.txt file
```{r}
activities<-read.table("UCI HAR Dataset/activity_labels.txt")
```
## Merging data

Read and merge activities of train and test set, which are in y_train.txt and y_test.txt files respectively by using rbind function
```{r}
activity.code<-rbind(read.table("UCI HAR Dataset/test/y_test.txt"), read.table("UCI HAR Dataset/train/y_train.txt"))
```
Join activities and activty.code data sets by code using merge function
```{r}
activity<-merge(activity.code, activities, by="V1")
```
Read and merge subjects of train and test set from subject_train.txt and subject_test.txt into one data set called subject by using rbind function
```{r}
subject<-rbind(read.table("UCI HAR Dataset/test/subject_test.txt"), read.table("UCI HAR Dataset/train/subject_train.txt"))
```
Read and merge measurements of train and test data sets into one data set called measurments
```{r}
measurements<-rbind(read.table("UCI HAR Dataset/test/X_test.txt"), read.table("UCI HAR Dataset/train/X_train.txt"))
```
Assign variable names to variables of measurements data set
```{r}
colnames(measurements)<-feature_names
```
Extract mean and std measurements from data set by subsetting
```{r}
mean.and.std.measurements<-measurements[mean.and.std]
```
Add the subject column to the measurments data set
```{r}
mean.and.std.measurements$subject <- subject$V1
```
Add activity column to the measurments data set
```{r}
mean.and.std.measurements$activity <- activity$V2
```
## Creating final data set

Create second data set  with the average of each variable for each activity and each subject using aggregate function
```{r}
aggr.by.activity.and.subject<-aggregate(mean.and.std.measurements[,!(names(mean.and.std.measurements) %in%c("activity", "subject"))], by=list(activity=mean.and.std.measurements$activity, subject=mean.and.std.measurements$subject), FUN=mean)
```

## Creating txt file with the final data set
write.matrix(aggr.by.activity.and.subject, file = "tidy data.txt", sep = "\t        ")